Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2012-10-24 documents/mixtures/mixture_models.tex
address
dstn comments

In addition to these, there are general S\'ersic (``ser'') profiles, of
which the exp and dev profiles are special cases.  The general ser
profile has one parameter (the ``index'') $n$:
\begin{eqnarray}\displaystyle
Q^{\ser}(\spos\given n) &\equiv& \exp(-\alpha^{\ser}(n)\,[|\spos|^{1/n} - 1])
\\
\{\alpha^{\ser}(2), \alpha^{\ser}(3), \alpha^{\ser}(5)\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
\quad ,
\end{eqnarray}
where we have given the constant $\alpha^{\ser}(n)$ for just a few
values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles given
above provide values for $n=1$ and $n=4$.)

@@

In \figurename~\ref{fig:ser}, we show the dependence of amplitudes
$a^{\ser}_m(n)$ and variances $v^{\ser}_m(n)$ on the ser index $n$.
There is clearly continuity; a valuable follow-up project would be to
give expressions for the amplitudes and variances as a function of ser
index $n$.  In the absence of cleverness our advice would be to make
use of smooth interpolation.

@@

Applying this PSF to an exp or dev galaxy is slightly more
complicated, because the galaxy has not just a flux $S_g$ and a
central position $\pos_g$; it also has a shape.  Because we are only
considering these simple galaxies, we are only permitting ellipsoidal
shapes, which can be represented by a semi-major axis $a$, a
semi-minor axis $b$, and a position angle $\phi$, or equivalently by
eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
an affine transformation $\affine_g$ that takes a circle to the
relevant ellipse (and is therefore a general representation of an
ellipse; it is also the matrix square root of the symmetric variance
tensor describing the ellipse).  The galaxy is distorted by this
affine transformation \emph{prior} to PSF convolution, so the
focal-plane image (PSF-convolved intensity field) for a general (say)
$\exp$ galaxy is given by
\begin{eqnarray}\displaystyle
I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,a^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
\\
\var_{gm} &\equiv& \affine_g\cdot\var^{\exp}_m\cdot\transpose{\affine_g}
\\
\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
\quad ,
\end{eqnarray}
where $a$ and $b$ are the major and minor axis lengths of the galaxy
ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
eigenvectors on the sky pointing in the major-axis and minor-axis
directions respectively.  Implicitly all vectors are two-dimensional
column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
matrix that contains the ``shape'' (position angle, major-axis, and
ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
$\lux$ cases are all similar.  Note the important and key result of
this \documentname, to wit, that a MoG galaxy model (with $M$
components) convolved with a MoG PSF model (with $K$ components)
yields a MoG model image (with $[M\,K]$ components).  In
\figurename~\ref{fig:example} we show how we are using these MoG
approximations in \project{The Tractor}---a generative modeling
framework for measuring astronomical objects---to render PSF-convolved
galaxy images.

@@

\clearpage
\begin{figure}
\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.pdf}%
\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.pdf}\\
\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.pdf}%
\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.pdf}
\caption{Three ser profiles---with $n=2$, $3$, and $5$---and
  approximations.  The top-left, top-right, and bottom-left panels are
  equivalent to those in the top-right of \figurename~\ref{fig:exp}.
  \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser}_m$
  and root-variances $\sqrt{v^{\ser}_m}$ on ser index $n$.\label{fig:ser}}
\end{figure}

>>>>

In addition to these, there are general S\'ersic (``ser'') profiles, of
which the exp and dev profiles are special cases.  The general ser
profile has one parameter (the ``index'') $n$:
\begin{eqnarray}\displaystyle
Q^{\ser(n)}(\spos) &\equiv& \exp(-\alpha^{\ser(n)}\,[|\spos|^{1/n} - 1])
\\
\{\alpha^{\ser(2)}, \alpha^{\ser(3)}, \alpha^{\ser(5)}\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
\quad ,
\end{eqnarray}
where we have given the constant $\alpha^{\ser(n)}$ for just a few
values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles given
above provide values for $n=1$ and $n=4$.)

@@

In \figurename~\ref{fig:ser}, we show the dependence of amplitudes
$a^{\ser(n)}_m$ and variances $v^{\ser(n)}_m$ on the ser index $n$.
There is clearly continuity; a valuable follow-up project would be to
give expressions for the amplitudes and variances as a function of ser
index $n$.  In the absence of cleverness our advice would be to make
use of smooth interpolation.

@@

Applying this PSF to an exp or dev galaxy is slightly more
complicated, because the galaxy has not just a flux $S_g$ and a
central position $\pos_g$; it also has a shape.  Because we are only
considering these simple galaxies, we are only permitting ellipsoidal
shapes, which can be represented by a semi-major axis $a$, a
semi-minor axis $b$, and a position angle $\phi$, or equivalently by
eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
an affine transformation $\affine_g$ that takes a circle to the
relevant ellipse (and is therefore a general representation of an
ellipse; it is also the matrix square root of the symmetric variance
tensor describing the ellipse).  The galaxy is distorted by this
affine transformation \emph{prior} to PSF convolution, so the
focal-plane image (PSF-convolved intensity field) for a general (say)
$\exp$ galaxy is given by
\begin{eqnarray}\displaystyle
I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,a^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
\\
\var_{gm} &\equiv& \affine_g\cdot\var^{\exp}_m\cdot\transpose{\affine_g}
\\
\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
\quad ,
\end{eqnarray}
where $a$ and $b$ are the major and minor axis lengths of the galaxy
ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
eigenvectors in image coordinates pointing in the major-axis and
minor-axis directions respectively.  Implicitly all vectors are
two-dimensional column vectors, and $\affine_g$ is a $2\times 2$
affine transformation matrix that contains the ``shape'' (position
angle, major-axis, and ellipticity) information about the galaxy.  The
$\dev$, $\luv$, and $\lux$ cases are all similar.  Note the important
and key result of this \documentname, to wit, that a MoG galaxy model
(with $M$ components) convolved with a MoG PSF model (with $K$
components) yields a MoG model image (with $[M\,K]$ components).  In
\figurename~\ref{fig:example} we show how we are using these MoG
approximations in \project{The Tractor}---a generative modeling
framework for measuring astronomical objects---to render PSF-convolved
galaxy images.

@@

\clearpage
\begin{figure}
\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.pdf}%
\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.pdf}\\
\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.pdf}%
\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.pdf}
\caption{Three ser profiles---with $n=2$, $3$, and $5$---and
  approximations.  The top-left, top-right, and bottom-left panels are
  equivalent to those in the top-right of \figurename~\ref{fig:exp}.
  \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser(n)}_m$
  and root-variances $\sqrt{v^{\ser(n)}_m}$ on ser index $n$.\label{fig:ser}}
\end{figure}
